---
title: "Part 2 - Cluster Annotation and Analysis"
output: html_document
---

# Annotating the cell types using Human Protein Atlas

### load necessary libraies

```{r}
library(Seurat)
library(tidyverse)
library(SingleR)
library(dplyr)
library(celldex)
library(RColorBrewer)
library(SingleCellExperiment)
```

### read in the 3k pbmc data

```{r}
# this returns a seurat object
pbmc<- readRDS("/Users/hawacoulibaly/Documents/GitHub/scHumanBcellFlu/data/pbmc_3k.RDS")
pbmc
```

### \### plot the UMAP and TSNE side by side tVisualize clusters

```{r}
p1<- DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 2 ,label.size = 8)
p2<- DimPlot(pbmc, reduction = "tsne", label = TRUE, pt.size = 2 ,label.size = 8)
p1+p2
```

### load the HPCA refrence dataset

```{r}
### load HPCA immune reference
hpca.ref <- celldex::HumanPrimaryCellAtlasData()
```

```{r}
# Let’s convert our Seurat object to single cell experiment (SCE) for convenience. After this, using SingleR becomes very easy:
sce <- as.SingleCellExperiment(DietSeurat(pbmc))
sce
```

Now, we can use the Human Protein Cell Atlas (HPCA) dataset for the 3k pmbc cells to annotate clusters.

```{r}
### Assign cell identity to clusters
hpca.main <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.main, clusters=pbmc@meta.data$seurat_clusters)
hpca.fine <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.fine, clusters=pbmc@meta.data$seurat_clusters)
print('DONE')

```

### View main assignment

```{r}
table(hpca.main$pruned.labels)
```

### View fine assignment

```{r}
table(hpca.fine$pruned.labels)
```

### View fine assignment

```{r}
table(hpca.fine$labels)

```

### save metadata

```{r}
# write.csv(hpca.main,"/Users/hawacoulibaly/Documents/GitHub/scHumanBcellFlu/output/hpca_cluster_annotation_main.csv" )
# write.csv(hpca.fine,"/Users/hawacoulibaly/Documents/GitHub/scHumanBcellFlu/output/hpca_cluster_annotation_fine.csv" )
print('DONE')

```

Finally, let’s visualize the cluster-wise fine-grained annotations.

```{r}
pbmc.main <- RenameIdents(pbmc, `0` = "B cell", `1` = "B cell", `2` = "B cell", `3` = "B cell", `4` = "B cell", `5` = "B cell")
pbmc.fine <-  RenameIdents(pbmc, `0` = "0 : Immature B cell", `1` = "1 : Immature B cell", `2` = "2 : Immature B cell", `3` = "3 : Memory B cell", `4` = "4 : Memory B cell", `5` = "5 : Memory B cell")
pbmc.final <-  RenameIdents(pbmc, `0` = "Immature B cell", `1` = "Immature B cell", `2` = "Immature B cell", `3` = "Memory B cell", `4` = "Memory B cell", `5` = "Memory B cell")


# p1 <- DimPlot(pbmc.main, label = T , reduction = 'umap', repel = T, label.size = 3) + NoLegend()
p2 <- DimPlot(pbmc.fine, label = T , reduction = 'umap', repel = T, label.size = 3) + NoLegend()
p3 <- DimPlot(pbmc.final, label = T , reduction = 'umap', repel = T, label.size = 3)

p2+p3
```

Let's validate using Immature and Memory B cell markers

```{r}
### Immature B cell markers
features = c( 'CD74', 'CD79A', 'CD83', 'HLA-DPB1', 'HLA-DRA', 'HLA-DRB1')
DotPlot(pbmc, features = features) + RotatedAxis()
```
```{r}
### Memory B cell markers
features = c('MS4A1','PTPRC','CD45RO')
DotPlot(pbmc, features = features) + RotatedAxis()
```


