---
title: "Part 2 - Cluster Annotation"
output: html_document
---

# Annotating the cell types using Human Protein Atlas

### load necessary libraries

```{r}
library(Seurat)
library(tidyverse)
library(SingleR)
library(dplyr)
library(celldex)
library(RColorBrewer)
library(SingleCellExperiment)
```

### read in the 3k pbmc data

```{r}
# this returns the seurat pre_processed object
pbmc<- readRDS("/Users/hawacoulibaly/Documents/GitHub/scHumanBcellFlu/data/pbmc_3k_pre_process.RDS")
pbmc
```

```{r}
# find number of cells for SCTransform
ncol(pbmc)
```

```{r}
pbmc <- NormalizeData(pbmc)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
```

SCTransform command replaces NormalizeData, ScaleData, and FindVariableFeatures. We will also correct for % MT genes and cell cycle scores using vars.to.regress variables percent MT, S AND G2M Scores

# First, let’s calculate cell cycle scores

```{r}
cc.genes.updated.2019
```

```{r}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes

pbmc <- CellCycleScoring(pbmc, s.features = s.genes, g2m.features = g2m.genes)
table(pbmc[[]]$Phase)
```

# Now, let's apply SCTransform

```{r}
pbmc <- SCTransform(pbmc, method = "glmGamPoi", ncells = 2456, variable.features.n = 2000 , vars.to.regress = c("percent.mt","S.Score","G2M.Score"), verbose = F)
pbmc
```

# After this let’s do standard PCA, UMAP, and clustering. the exact number can be adjusted depending on your dataset.

```{r}
# B-cell markers
features <- c('Leu-12' , 'CD27' , 'IGM' , 'CD19' , 'ZAP70' , 'IGD' , 'CD38' , 'CD45' , 'CD24' , 'CD21' , 'CD20' , 'CD5' , 'CD74' , 'MS4A1' , 'CD79A' , 'CD79B' , 'BLNK' , 'CD37' , 'CD11C' , 'CD95' , 'T-bet' , 'MZB1' , 'LILRA4' , 'IGJ' , 'CD79' , 'CD22' , 'HLA-A' , 'IGHG2' , 'CALR' , 'TNFRSF17' , 'JCHAIN' , 'XBP1' , 'BACH2' , 'CD127' , 'CD25' , 'CD1C' , 'CR2' , 'CD81' , 'CD40' , 'CD72' , 'FCGR2A' , 'IGHM' , 'PAX5' , 'SPIB' , 'ITGB1' , 'FCRLA' , 'IGHA1' , 'IGHD' , 'FCER2' , 'TCL1A' , 'BCL11A' , 'PLPP5' , 'FCRL1' , 'TNFRSF13C' , 'YBX3' , 'AHNAK' , 'HLA-DPB1' , 'NFKBIA' , 'SWAP70' , 'AIM2' , 'IL4R' , 'IGKC')

pbmc <- RunPCA(pbmc, features = features, verbose = F)
pbmc <- RunUMAP(pbmc, dims = 1:10, verbose = F)
pbmc <- RunTSNE(pbmc, dims = 1:10, verbose = F)

pbmc <- FindNeighbors(pbmc, features = features, verbose = F)
pbmc <- FindClusters(pbmc, verbose = F, resolution = 0.5)

table(pbmc[[]]$seurat_clusters)

```

### \### plot the UMAP and TSNE side by side to visualize clusters

```{r}
p1<- UMAPPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 2 ,label.size = 6)
p2<- TSNEPlot(pbmc, reduction = "tsne", label = TRUE, pt.size = 2 ,label.size = 6)
p1+p2
```

### load the HPCA reference dataset

```{r}
### load HPCA immune reference
hpca.ref <- celldex::HumanPrimaryCellAtlasData()
```

```{r}
# Let’s convert our Seurat object to single cell experiment (SCE) for convenience. After this, using SingleR becomes very easy:
sce <- as.SingleCellExperiment(DietSeurat(pbmc))
sce
```

Now, we can use the Human Protein Cell Atlas (HPCA) dataset for the 3k pmbc cells to annotate clusters.

```{r}
### Assign identity to clusters
hpca.fine <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.fine, clusters=pbmc@meta.data$seurat_clusters)
hpca.ont <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.ont, clusters=pbmc@meta.data$seurat_clusters)
print('DONE')

```

### View fine assignment

```{r}
table(hpca.fine$labels)
```

CLUSTER LABELS

0 B_cell:immature 1 B_cell:immature 2 B_cell:Memory 3 B_cell:immature 4 B_cell:Memory 5 B_cell:immature 6 B_cell:immature

### View ontology assignment

```{r}
table(hpca.ont$pruned.labels)
```

CL:0000787 - memory B cell CL:0000816 - immature B cell

CLUSTER LABELS

0 memory B cell 1 immature B cell 2 memory B cell 3 memory B cell 4 memory B cell 5 memory B cell 6 memory B cell

### save metadata

```{r}
#write.csv(hpca.fine,"hpca_Bcell_cluster_annotation_fine.csv" )
#write.csv(hpca.ont,"hpca_B_cell_cluster_annotation_ont.csv" )

print('DONE')
```

Finally, let’s visualize the cluster-wise fine-grained annotations.

```{r}
pbmc.ont <-  RenameIdents(pbmc, `0` = "0 : Memory B cell", `1` = "1 : Immature B cell", `2` = "2 : Memory B cell", `3` = "3 : Memory B cell", `4` = "4 : Memory B cell", `5` = "5 : Memory B cell", `6` = "6 : Memory B cell")
pbmc.final <-  RenameIdents(pbmc, `0` = "Memory B cell", `1` = "Immature B cell", `2` = "Memory B cell", `3` = "Memory B cell", `4` = "Memory B cell", `5` = "Memory B cell", `6` = "Memory B cell")
```

```{r}
UMAPPlot(pbmc.ont , reduction = 'umap' ,label = TRUE, pt.size = 2 ,label.size = 4)
```

visualize the B cell subtypes

```{r}
p1 <- UMAPPlot(pbmc.final, reduction = 'umap',  label = TRUE, pt.size = 2 ,label.size = 4)
p2 <- TSNEPlot(pbmc.final, reduction = "tsne", label = TRUE, pt.size = 2 ,label.size = 4)

p1+p2
```

```{r}
DimPlot(object = pbmc.final, reduction = 'umap' , split.by = 'ident')
```

```{r}
DimPlot(object = pbmc.final, reduction = 'tsne' , split.by = 'ident')
```

Our analysis leads us to the conclusion that the dataset is mostly constituted of MEMORY B CELLS

For further validation, let's explore using canonical Memory B cell markers <https://www.rndsystems.com/research-area/b-cell-markers>

B220/CD45R ,C1qR1/CD93, CD19, CD21, CD27/TNFRSF7, CD40/TNFRSF5, CD20, MHC class II (I-A/I-E), TACI/TNFRSF13B AIM2, CD20 , CD27, CD79A, IGD, JCHAIN, MS4A1

```{r}
features <- c('CD19', 'CD27', 'CD40' , 'CD79A' ,'MS4A1' )

DotPlot(pbmc.final, features = features, cols=c("#5F4B8BFF", "#ED2B33FF"), assay = "RNA", col.min = 0.3, col.max = 0.8, dot.min=0.12, dot.scale = 8, cluster.idents=F) + RotatedAxis()
```

feature plot of Memory B cell markers

```{r}
FeaturePlot(object = pbmc.final, features = "MS4A1", pt.size = 2,  split.by = 'ident', keep.scale = 'feature')
```

Immature B cell markers retrieved from the CellMarker 2.0 databasee <https://doi.org/10.1093/nar/gkac947> <http://117.50.127.228/CellMarker/CellMarkerBrowse.jsp>

Species Tissue Class Tissue Type Cancer Cell name Cell marker Source Supports Detail Human Blood Peripheral blood Normal cell Immature B cell CD10 Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell CD74 Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell CD79A Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell CD83 Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell HLA-DPB1 Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell HLA-DRA Experiment 1 DETAIL Human Blood Peripheral blood Normal cell Immature B cell HLA-DRB1 Experiment 1 DETAIL

```{r}
features <- c('CD74' , 'CD79A' , 'CD83' , 'HLA-DPB1' , 'HLA-DRA' , 'HLA-DRB1')

DotPlot(pbmc.final, features = features, cols=c("#5F4B8BFF", "#ED2B33FF"), assay = "RNA", col.min = 0.3, col.max = 0.8, dot.min=0.12, dot.scale = 8,
        cluster.idents=F) + RotatedAxis()
```

feature plot of Immature B cell markers

```{r}
FeaturePlot(object = pbmc.final, features = "CD74", pt.size = 2,  split.by = 'ident', keep.scale = 'feature')
```
### save object
```{r}
# saveRDS(pbmc.final, "/Users/hawacoulibaly/Documents/GitHub/scHumanBcellFlu/data/pbmc_3k_annotated.RDS")
```

